
---
# Keep YAML above the first chunk.
title: "<span style='font-size: 34px'>DARE-Newcrest data challenge - flotation plant dataset</span>"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

library(knitr)          # To set chunk options.

# Global knitr options.
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

DARE 2022

In this R Markdown file, we will provide a template for the DARE-Newcrest data challenge. The main steps covered here will be:

- Load the data
- Provide an overview of what is in the data
- Provide an example of a terribly performing baseline model
- Provide functions to quantify model predictive performance

### Imports and settings
Everything we need to get started.

```{r chunk_one}

source("functions_R/R_data_functions.R", local = knitr::knit_global())
source("functions_R/R_helper_functions.R", local = knitr::knit_global())
source("functions_R/R_plotting_functions.R", local = knitr::knit_global())

```

### Load the data
Briefly look at the contents and structure of the data.

```{r chunk_two}

process_data <- load_raw_data()
head(process_data)

```

```{r chunk_three}

```
